\sysfull

### 2018-04-17 paper/background.tex

Prior systems, including the task-based models, suffer from two major overheads:
{\em frequent non-volatile memory access} and {\em incorrectly size tasks}.

@@

\textbf{Frequent Non-volatile Memory Access.} 
Previous systems frequently access the non-volatile memory directly, which we claim is a major source of inefficiency.
Checkpoint-based models copy the entire volatile state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile
memory on every checkpoint.
Task-based models, which do not take checkpoints, still have to copy a subset of the non-volatile data to maintain memory consistent in WAR dependence (e.g., redo-logging).
Table~\ref{table:chechpoint_comparison} summarizes the data that gets copies to and from the non-volatile memory for memory consistency and forward progress.
Ratchet~\cite{ratchet} and Clank~\cite{hicks_isca_2017}, which uses a whole non-volatile memory, obviously also has a high number of non-volatile memory access
for normal operations in addition to taking checkpoints.

%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.  
% We do not need to emphasize the following figure! (Sinan)
%\begin{figure}
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
%	\caption{Cost of \emph{write} operation}
%\end{subfigure}%
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
%	\caption{Cost of \emph{read} operation}
%\end{subfigure}
%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
%\end{figure}
To understand the expense of non-volatile memory access, we measured the energy consumption of volatile (SRAM) and non-volatile (FRAM) memory 
operations for the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor voltage drop on a capacitor powering the device 
across 1600 read and write accesses to each memory type, randomizing to avoid caching. Our experiment showed that a non-volatile memory access 
consumes around 1.5 times the energy of a volatile access. The result shows that non-volatile memory access is more expensive than the volatile
memory access, and prior works with frequent non-volatile memory accesses is inefficient.
%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. This observation is also consistent with the TI specification~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek} 
Based on this observation, one of the primary motivation of \sys is to promote accessing SRAM more frequently rather than FRAM in order to 
reduce memory access energy overhead and to increase the time between power failures.

@@

\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or programmer-defined tasks can both be incorrect or inefficient.
If the code between checkpoints or the task (for brevity, we will refer to both as ''tasks'' in this section) consumes {\em more} 
energy than the device can buffer, the task will never complete 
using buffered energy {\em preventing forward progress}, making the system incorrect. 
%
If the task consumes far {\em less} energy than a device can buffer, the system may operate {\em inefficiently},
checkpointing or transitioning between tasks too often than needed. 
%
Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is challenging both to the
programmer and the compiler, since estimating the exact energy use of an arbitrary code is complicated.
When {\em porting} apps from one device to another, the problem gets much worse because
an excessively costly task on one device may be a relatively short task on a device with a larger buffer, and vice versa. 
Buffer sizes of different platforms vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
complicating the problem.
%
As a solution, \sys's execution model presents {\em task coalescing}, which merges multiple tasks to amortize the overhead
when tasks are too small, and {\em task downscaling}, which breaks up a task when the task is too large. 
%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity. Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer. The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead. \sys's execution model based on {\em task coalescing} and {\em task downscaling}.  Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately. Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.

>>>>

Prior systems, including the task-based models, suffer from two major overheads:
{\em frequent non-volatile memory access} and {\em incorrectly sized, inefficient or non-terminating tasks}.

@@

\textbf{Frequent Non-volatile Memory Access.} 
Previous systems frequently access the non-volatile memory directly, which 
is a potential source of inefficiency.  Checkpointing systems copy all volatile
state~\cite{dino, mementos, ratchet, hicks_isca_2017} to non-volatile memory at
every checkpoint.  Task-based models, which do not take checkpoints, back up a
subset of non-volatile data to maintain memory consistency using, e.g.,
redo-logging.  Table~\ref{table:chechpoint_comparison} summarizes the data that
gets copies to and from the non-volatile memory in service of memory
consistency and progress preservation.  Ratchet~\cite{ratchet} and
Clank~\cite{hicks_isca_2017}, which require all memory to be non-volatile
perform high frequency non-volatile memory accesses, in to checkpointing
frequently.

%Intermittent execution models incur overheads to checkpoint data~\cite{dino,ratchet,quickrecall,mementos}, manage channels~\cite{chain}, or privatize and commit data~\cite{alpaca}. Table~\ref{table:chechpoint_comparison} compares overheads for recent intermittent execution models (cf. \cite[Sec. 2.4]{alpaca}.) The mechanism responsible for overhead in each model varies, but the bulk of overhead in all models is a manipulating non-volatile memory. Checkpointing moves data to and from non-volatile memory, channel accesses manipulate non-volatile data~\cite{chain}, privatization copies from and commits to non-volatile memory~\cite{alpaca}, and idempotence solutions~\cite{ratchet} use {\em only} non-volatile memory.  
% We do not need to emphasize the following figure! (Sinan)
%\begin{figure}
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_write}
%	\caption{Cost of \emph{write} operation}
%\end{subfigure}%
%\begin{subfigure}[t]{.49\columnwidth}
%	\centering \includegraphics[width=\columnwidth]{figures/fram_read}
%	\caption{Cost of \emph{read} operation}
%\end{subfigure}
%	\caption{\textbf{Cost of accessing volatile (SRAM)/non-volatile (FRAM) memory during write/read operation.} \textcolor{red}{figures need to be merged into one}}\label{fig:framEnergy}
%\end{figure}
We measured the energy cost of volatile SRAM and non-volatile FRAM accesses on
the MSP430FR5969~\cite{msp430datasheet} MCU. We used EDB~\cite{edb} to monitor
voltage drop on a capacitor powering the device across 1600 read and write
accesses to each memory, randomizing accesses to avoid caching effects. Our
data showed that a non-volatile memory access consumes around 1.5 times the
energy of a volatile access. The data confirm that, because non-volatile memory
accesses are more expensive than volatile memory accesses, prior systems that
frequently access non-volatile memory may be artificially inefficient.
%~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek}. 
% Figure~\ref{fig:framEnergy} summarizes the data, showing that a non-volatile memory access consumes around 1.5 times the energy of a volatile access in this device. This observation is also consistent with the TI specification~\cite{}. \todo{add citation to TI specification on FRAM energy cost}{Przemek} 
\sys is motivated by this observation: \sys 
moves data pages from FRAM to SRAM to maximize SRAM accesses and minimize FRAM
accesses, to decrease energy consumption and increase the time between power
failures.

@@

\textbf{Incorrectly Sized Tasks.} Compiler-inserted checkpoints or
programmer-defined tasks can both both incorrect and non-terminating or
inefficient.  If code between consecutive checkpoints or in a task (we refer to
both scenarios as ''code in a task'' in this section) consumes {\em more}
energy than the fixed, maximum energy that the device can buffer, then the task
will never be able to complete using buffered energy.  Such a task is
non-terminating, {\em prevents forward progress}, and makes the program
incorrect. 
%
If the task consumes far {\em less} energy than a device can buffer, the system
may operate {\em inefficiently}, checkpointing or transitioning between tasks
more often than needed. 
%
Avoiding excessively costly, non-terminating tasks and short, high-overhead
tasks is challenging both to the programmer and the compiler, because
estimating the exact energy use of an arbitrary code is complicated.  When {\em
porting} a program from one platform to another, the problem becomes worse
because an excessively costly task on one device may be a relatively short task
on a device with a larger buffer.  Buffer sizes on different platforms vary
widely from, e.g., 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo},
complicating the problem.
%
As a solution, \sys's execution model presents {\em task coalescing}, which
merges multiple tasks to amortize the overhead when tasks are too small, and
{\em task downscaling}, which breaks up a task when the task is too large to complete. 
%When a task and its successor both complete, the two tasks were interleaved by a task boundary that preserves progress and state (i.e., checkpoint or commit). However, absent a power failure, the work to preserve state was {\em unnecessary}, incurring overhead on the tasks' executions. Avoiding excessively costly, non-terminating tasks and short, high-overhead tasks is a programming challenge, given fixed hardware (no support from dedicated circuitry to measure the state of energy buffer) with a fixed energy buffering capacity. Buffer sizes from prior work on intermittently-powered computing hardware vary widely from 20\,$\mu $F~\cite{rodriguez_tbcs_2015} to 0.1\,F~\cite{moo}. Sizing tasks complicates {\em porting} code from one device to another. An excessively costly task on a device with a small energy buffer may be a relatively short task on a device with a larger buffer. The key problem is that using existing systems, by defining tasks statically for a fixed energy buffer, the programmer either risks non-termination or poor performance due to high overhead. \sys's execution model based on {\em task coalescing} and {\em task downscaling}.  Coalescing dynamically merges multiple tasks based on execution behavior to amortize the cost of committing each task separately. Downscaling breaks up a single task into multiple subtasks that commit individually to avoid executing a single task that is too large to complete using the device's buffered energy.  Section~\ref{sec:task_adaptation}describes \sys's adaptive coalescing and downscaling mechanisms in detail. The main enabler for \sys's adaptation is its novel \emph{memory virtualization} mechanism, which is described in Section~\ref{sec:memory_virtulaization}.
